Causal Bandits: Learning Good Interventions via Causal Inference

### 2016-02-05 causal-bandit/icml-2015/discussion.tex

\paragraph{Using the Reward Signal}

\paragraph{Cumulative Regret}
Although we have focused on simple regret in our analysis, a natural but open question is whether cumulative regret bounds can be obtained for the general causal bandit problem we have proposed.
In the case of the parallel bandit problem we can slightly modify the analysis from \citep{wu2015online} on bandits with graph feedback to get optimal cumulative regret guarantees.
Their model assumes the rewards for all arms $a$ are Gaussian with mean $\mu_a$ and variance $\sigma^2_{ab}$ and that playing an arm $a$ will reveal a side observation $Y_{ab}$ of the reward for all arms $b$ distributed with mean $\mu_b$ and variance $\sigma^2_{ab}$.
We can build a similar dependence structure with variances for a Bernoulli reward variable that is derived from the $\vec{q}$ vector of probabilities.
Even though the original results are for Gaussian rewards we believe the analysis will go through largely unchanged.

Exploiting the variance structure in this way seems necessary even in the parallel bandit case.
If instead we choose to use other graph feedback models for cumulative regret we would require those with time varying graphs~\citep{Alon2015,Kocak2014}.
Results for these models give a cumulative regret of $\bigo{\sqrt{\bar{\alpha} T}}$ where $\bar{\alpha}$ is the average independence number of the time varying graphs.
On each round, the feedback graph over the $2N$ possible interventions over $N$ binary variables would connect the chosen intervention to the corresponding intervention for each of the other $N-1$ observed variables.
Unfortunately, in this case the resulting star graphs have an $\bar{\alpha} = N$ and so the cumulative regret bounds would improve over those obtained by ignoring the causal graph structure.

% In our algorithm, we have only used the side information provided by the $do()$ action about other actions. Since the $do()$ action fully reveals the value of alternate actions we could have incorporated this information via the graph feedback model \cite{Mannor2011}, where at each timestep the feedback graph $G_t$ is selected stochastically, dependent on $\boldsymbol{q}$, and revealed after an action has been chosen. The feedback graph is distinct from the causal graph. A link $A \rightarrow B$ in $G_t$ indicates that selecting the action $A$ reveals the reward for action $B$. For this specific problem, $G_t$ will always be a star graph with the action $do()$ connected to half the remaining actions. The Exp3-IX algorithm \cite{Kocak2014} was developed for the adversarial version of this problem and has regret $\bigo{\sqrt{\bar{\alpha}T}}$, where $\bar{\alpha}$ is the average independence number of $G_t$. In our case $\bar{\alpha} = \frac{N}{2}$ so we again obtain the regret of the standard bandit algorithm. The issue here is that a malicious adversary can select the same graph each time, such that the rewards for half the arms are never revealed by the informative action. This is equivalent to a, nominally, stochastic selection of feedback graph where $\boldsymbol{q} = \boldsymbol{0}$

@@

% Non-observable variables? 
% Currently assume all P_a are available (implied by observablity)
%   - do calculus' identifability
%   - non-ident variables into infinite variance case.

@@

\paragraph{Partially or Completely Unknown Causal Graph}
A much more difficult generalisation would be to consider causal bandit problems where the causal graph is completely unknown or known to be a member of class of models.
The latter case arises naturally if we assume free access to a large observational dataset, from which the Markov equivalence class can be found via causal discovery techniques. 
Work on the problem of selecting experiments to discover the correct causal graph from within a Markov equivalence class~\cite{Eberhardt2005,eberhardt2010causal,hauser2014two,Hu2014} could potentially be incorporated into a causal bandit algorithm as part of a pre-processing step.
\todom{Ideas? Relevant refs?}
Simultaneously learning a completely unknown causal model while estimating the rewards of interventions would be much more challenging.

>>>>

\paragraph{Making Better Use of the Reward Signal}
Several existing algorithms for best arm identification are based on ``successive rejection'' (SR) of arms based on UCB-like bounds on their rewards~\citep{Even-Dar2002}.
In contrast, our algorithms completely ignore the reward signal when developing their arm sampling policies and only use the rewards when estimating $\hat{\mu}_a$.
Incorporating the reward signal into our sampling techniques or designing more adaptive reward estimators that focus on high reward interventions is an obvious next step.
This would likely improve the poor performance of our causal algorithm relative to the sucessive rejects algorithm for large $m$, as seen in Figure~\ref{fig:simple_vs_m}.

\paragraph{Cumulative Regret}
Although we have focused on simple regret in our analysis, a natural but open question is whether cumulative regret bounds can be obtained for the general causal bandit problem we have proposed.
In the case of the parallel bandit problem we can slightly modify the analysis from \citep{wu2015online} on bandits with graph feedback to get optimal cumulative regret guarantees.
Their model assumes the rewards for all arms $a$ are Gaussian with mean $\mu_a$ and variance $\sigma^2_{ab}$ and that playing an arm $a$ will reveal a side observation $Y_{ab}$ of the reward for all arms $b$ distributed with mean $\mu_b$ and variance $\sigma^2_{ab}$.
We can build a similar dependence structure with variances for a Bernoulli reward variable that is derived from the $\vec{q}$ vector of probabilities.
\todom{Check this!}
Even though the original results are for Gaussian rewards we believe the analysis will go through largely unchanged.

Exploiting the variance structure in this way seems necessary even in the parallel bandit case.
If instead we choose to use other graph feedback models we would require those with time varying graphs~\citep{Alon2015,Kocak2014}.
Results for these models give a cumulative regret of $\bigo{\sqrt{\bar{\alpha} T}}$ where $\bar{\alpha}$ is the average independence number of the time varying graphs.
On each round, the feedback graph over the $2N$ possible interventions over $N$ binary variables would connect the chosen intervention to the corresponding intervention for each of the other $N-1$ observed variables.
Unfortunately, in this case the resulting star graphs have an $\bar{\alpha} = N$ and so the cumulative regret bounds would improve over those obtained by ignoring the causal graph structure.

@@

@@

\paragraph{Partially or Completely Unknown Causal Graph}
A much more difficult generalisation would be to consider causal bandit problems where the causal graph is completely unknown or known to be a member of class of models.
The latter case arises naturally if we assume free access to a large observational dataset, from which the Markov equivalence class can be found via causal discovery techniques. 
Work on the problem of selecting experiments to discover the correct causal graph from within a Markov equivalence class~\cite{Eberhardt2005,eberhardt2010causal,hauser2014two,Hu2014} could potentially be incorporated into a causal bandit algorithm as part of a pre-processing step.
Simultaneously learning a completely unknown causal model while estimating the rewards of interventions would be much more challenging.
\citet{Hu2014b} show that ...

### 2016-02-05 causal-bandit/icml-2015/general.tex
Tweaks

Let $a \in \calA$ be an intervention and define random variable
\eq{
Z_a(X) = R_a(X) Y \ind{R_a(X) \leq B_a}\,,
}
where $B_a \geq 0$ is some constant to be chosen subsequently and
\eq{
R_a(X) = \frac{\P{\parents{Y}(X)|a}}{\Q{\parents{Y}(X)}}\,.
}
Abusing notation we define $m(\eta)$ by
\eq{
m(\eta) = \max_{a \in \calA} \EE_{\Ps_a}\left[\frac{\Pn{a}{\parents{Y}(X)}}{\Q{\parents{Y}(X)}}\right]\,.
}
We will shortly see that $m(\eta)$ is a measure of the difficulty of the problem that approximately cooincides 
with the version given in the previous section, which justifies the name overloading.

\begin{algorithm}[H]
\caption{General Algorithm}\label{alg:general}
\begin{algorithmic}
\STATE {\bf Input:} $T$, $\eta \in [0,1]^{|\calA|}$, $B \in [0,\infty)^{|\calA|}$
\FOR{$t \in \set{1,\ldots,T}$}
\STATE Sample action $a_t = a$ with probability $\eta$
\STATE Do action $a_t$ and observe $X_t$ and $Y_t$
\ENDFOR
\STATE For each $a \in \calA$ compute an estimate of its return:
\eq{
\forall a\in \calA \qquad \hat \mu_a = \frac{1}{T} \sum_{t=1}^T Z_a(X_t)
}
\STATE {\bf return} $\hat a^* = \argmax_a \hat \mu_a$
\end{algorithmic}
\end{algorithm}

>>>>

Let $a \in \calA$ be an intervention and define random variable
\eq{
Z_a(X) = R_a(X) Y \ind{R_a(X) \leq B_a}\,,
}
where $B_a \geq 0$ is some constant to be chosen subsequently and
\todom{Symbol clash with regret}
\eq{
R_a(X) = \frac{\Pn{a}{\parents{Y}(X)}}{\Q{\parents{Y}(X)}}\,.
}
Abusing notation we define $m(\eta)$ by
\eq{
m(\eta) = \max_{a \in \calA} \EE_{\Ps_a}\left[\frac{\Pn{a}{\parents{Y}(X)}}{\Q{\parents{Y}(X)}}\right]\,.
}
We will shortly see that $m(\eta)$ is a measure of the difficulty of the problem that approximately coincides with the version for parallel bandits, justifying the name overloading.

\todom{Change to $[0,1]^{\calA}$}
\begin{algorithm}[H]
\caption{General Algorithm}\label{alg:general}
\begin{algorithmic}
\STATE {\bf Input:} $T$, $\eta \in [0,1]^{|\calA|}$, $B \in [0,\infty)^{|\calA|}$
\FOR{$t \in \set{1,\ldots,T}$}
\STATE Sample action $a_t = a$ with probability $\eta$
\STATE Do action $a_t$ and observe $X_t$ and $Y_t$
\ENDFOR
\STATE For each $a \in \calA$ compute an estimate of its return:
\eq{
\forall a\in \calA \qquad \hat \mu_a = \frac{1}{T} \sum_{t=1}^T Z_a(X_t)
}
\STATE {\bf return} $\hat a^* = \argmax_a \hat \mu_a$
\end{algorithmic}
\end{algorithm}
